% skal handle om motivasjon bak ANN

Computers are specialized for calculating through ecplicit algorithms. Here the action is expressed almost as mathematical functions with an output value based on its input. 
The functions are expressed 

JAJA : SKRIV INTRO seinare. No over til kva eit ANN er:

In terms from mathematics a neural network can be called 
Since a neural network in a biological being is wery complex, with recurrent connections of adaptive strenght, %XXX KVA gjør det vanskelig å vurdere? XXX
, we will start from the buttom. 

The connections between the nodes of the neural network

In the simples form of neural networks, a network consisting of two nodes with a connection between, the neural network 



-------------------------------------------

In order to describe a neural network, I first need to establish the basic building blocks. The nodes of the network are neurons. The connections between neurons are called synapses.

The neuron has an internal state given by the voltage difference between the intracellular fluid and the extracellular fluid to the neuron. This is given by the ionic balance over the cellular membrane. 
If this value excedes a threshold the neuron will initiate an action potential.  

------------------------


The nodes of a neural network is nonlinear units with a boolean signal (an action potential). 
When the action potential reaches an output connection to an other neuron (a synapse) a signal is transmitted based on the trength of the connection. 

When this boolean signal reaches the output connections (synapses) of the neuron, a signal 


------------------------------------------------
The neuron:
The biological neuron is a node in the neural network with a value based on the voltage across the neurons cell membrane. If this value excedes a threshold, an action potential will be initiated. 
An action potential is a boolean signal that will spread along the axon of the neuron. 
This enshures that the signal transmitted through synapses will be invariant with the distance from where the action potential is initiated to the synapse.

When the action potential reaches the synapse, neurotransmitters are released from the presynaptic membrane of the synapse. 
These neurotransmitters will exite receptors located on the postsynaptic membrane, and the postsynaptic membranes will change value. This value will further change the value for the whole neuron.

On the presynaptic side of the synapse is the axon terminal of the presynaptic neuron that recieves the action potential generated as a function of the presynaptic neurons value. 
On the postsynaptic side lies a dendrite (input) of the postsynaptic neuron. The dendrites main function is to be a leaky integrator for the 

The synaptic connection between two neurons is directional, that is

Even though the signal transmitted through the neuron is a boolean signal, the transmission to other neurons are based on the synaptic connection to the next neuron. This connection will 


