\section{ANN, mi implementering.}

Før eg kan skrive om tid må eg skrive litt om mi implementering. Arbeidslister, osv. Først må eg lage dette...

Kladd til kva eg må hugse (skrevet mens eg skriver neste avsnitt, tid):
\begin{itemize}
	\item Snakke en del om at mitt ANN er simulert asynkront.
\end{itemize}



\section{Time}

Following the fact that the ANN will be simulated asynchronous, the computer resources will deside the size of the time iterations. The maximum size is desided by the real--time requirement of the task.

To make the software general, the simulated time should therefore be unconnected to the world time outside the simulation. To achieve this, a scheduler has been devised for my simulation.


In object oriented programming languages, we can make a linked list of elements. 
If the data of the elements are pointers, the pointers can be pointers to any kind of data. Let the list be of type \emph{$std::list<father*> LIST$} and \emph{$father$} contain virtual fuctions. 

When it is time for the scheduler-thread to start a new task, it calls \emph{LIST.front()$\rightarrow$doTask()}. 
Since \emph{doTask()} is overloaded in the derived classes, we can have different tasks assigned to each class derived from \emph{$<father>$}.




If we have a list of scheduled tasks $L = [a, b, c, T]$, all tasks derived from a common \emph{$<father>$}. 
All tasks can generate new tasks according to the rules of the derived classes, and when a task is done it will be removed from the list.

If we allways pick the first task, the next task to be executed will be task $a$. %All tasks can possibly generate new tasks.
If a generates new tasks $x_{a_1}$, this new task will be added at the end of the list.

\begin{equation}
	L = [b, c, T, x_{a_1}]
\end{equation}

Lets say that $b$ does not generate any new tasks and $c$ gives us two new tasks $[x_{c_1}, x_{c_2}]$, we get the list:
\begin{equation}
	L = [T, x_{a_1}, x_{c_1}, x_{c_2} ]
\end{equation}

As the syntax implies, the element $T$ is a rather unique task. $T$ is a const--instance of the timeIterator--class.


\emph{Skrevet dårlig herretter. Skrevet seint på kvelden:} %XXX XXX

\emph{T$\rightarrow$ doTask()} will have the responsibility for any tasks that have anything with time to do. 

First, it will the iterate the global time variable \emph{unsigned long timeIterations}.

Finally it will append a \emph{self--pointer} to the end of the task list.
\begin{equation}
	L = [x_{a_1}, x_{c_1}, x_{c_2}, T ]
\end{equation}

